---
title: "load_ffr"
subtitle: "This script loads Free-flowing-river data and merges it with HydroRIVERS"
format: html
---
# Import libraries
```{r}
library(janitor)
library(sf)
```

# Load Data
```{r}
# HydroRIVERS global
HydroRIVERS <- read_sf("/capstone/netzerohydro/data/raw/HydroRIVERS_v10_shp/HydroRIVERS_v10_shp") %>% 
  clean_names() # IMPORTANT: load shp not gdb

# Free flowing rivers global
ffr <- read_sf("/capstone/netzerohydro/data/raw/FFR_river_network.gdb") %>%   
  clean_names()
```

FFR has MULTILINESTRING Shape whereas HydroRIVERS has LINESTRING geometry. Run a simple check on a sample to see if FFR actually has disconnected vectors. 

```{r}
{
  check <- ffr %>%
    slice_sample(n = 100) %>% # Random sample
    mutate(n_segments = map_int(Shape, ~length(.x))) %>% # Count number of line segments in each MULTILINESTRING geometry
    summarise( # count features with multiple segments, find max segments
      multi_segment = sum(n_segments > 1), # Count how many have >1 segment
      max_segments = max(n_segments) # Find maximum number of segments
    )
  
  if (check$multi_segment > 0) {
    print(paste("WARNING:", check$multi_segment, "MULTILINESTRINGs have", 
                check$max_segments, "disconnected parts"))
  } else {
    print("All checked features are single segments (safe to convert to LINESTRING)")
  }
  
  rm(check)
}

# It does not, still if you plan to use the ffr geometry alongside HydroRIVERS you need to convert to linestring
ffr_linestring <- ffr %>%
  st_cast("LINESTRING")

```
This Warning:

Warning in st_cast.sf(., "LINESTRING") :
repeating attributes for all sub-geometries for which they may not be constant
 
Is harmless because ffr_linestring has the same number of rows as ffr. So nothing was repeated.
# Join HydroRIVERS and ffr 

```{r}
# Join FFR attributes to HydroRIVERS (Keeps HydroRIVERS CRS) 
HydroRIVERS_ffr <- HydroRIVERS %>%
  left_join(ffr %>% 
            st_drop_geometry(), # Remove ffr$geometry column (its the same as HydroRIVERS$Shape)
            by = c("hyriv_id" = "reach_id"))
```
 

# Save all three as RDS

```{r}
# Save HydroRIVERS
saveRDS(HydroRIVERS, "../../../../../capstone/netzerohydro/data/cleaned_v1/world_rivers.rds")

# Save FFR
saveRDS(ffr, "../../../../../capstone/netzerohydro/data/cleaned_v1/ffr.rds")

# Save joined object ffr + HydroRIVERS
saveRDS(HydroRIVERS_ffr, "../../../../../capstone/netzerohydro/data/cleaned_v1/world_rivers_ffr.rds")
```

