---
title: "0 — Environment, packages, and data loading"
subtitle: "Run this notebook first"
format: html
---

## Prerequisites

- **Libraries:** sf, janitor, dplyr, sfnetworks, tidygraph, igraph, readr, leaflet
- **Raw data:** edit `paths` below to match your machine
- **R session:** versions in README "Session info"

## Creates in your R session

`paths`, `target_crs`, `rivers`, `ffr`, `world_rivers_ffr`, `current_dams_gdw`, `future_dams_sf`

# Overview

Run this notebook once per session before notebooks 1–3. It loads packages, reads raw spatial data, and leaves named objects in your **global environment** (no intermediate `.rds` bundle).

Edit the `paths` list below if your data are not under `/capstone/netzerohydro/data`. More detail on datasets: README and Bren technical documentation.

# Paths

Change `data_root` or any individual path as needed.

```{r}
#| label: paths
#| message: false

data_root <- "/capstone/netzerohydro/data"

paths <- list(
  data_root = data_root,
  raw = file.path(data_root, "raw"),
  processed = file.path(data_root, "processed_data"),
  hydrorivers = file.path(data_root, "raw/HydroRIVERS_v10_shp/HydroRIVERS_v10_shp"),
  ffr_gdb = file.path(data_root, "raw/FFR_river_network.gdb"),
  gdw_snapped = file.path(data_root, "GDW_prj_1000m_from_StrOrd3_SNAPPED.gpkg"),
  fhred_snapped = file.path(data_root, "FhRED_prj_1000m_from_StrOrd3_SNAPPED.gpkg"),
  net_with_dams_rds = file.path(data_root, "processed_data/net_with_dams_GIS_v2.rds")
)

target_crs <- sf::st_crs(3857) # EPSG:3857 for the full analysis
dir.create(paths$processed, recursive = TRUE, showWarnings = FALSE)
```

# Packages

```{r}
#| label: packages
#| message: false
#| warning: false

library(sf)
library(sfnetworks)
library(tidygraph)
library(igraph)
library(dplyr)
library(janitor)
library(readr)
library(leaflet)
```

# Datasets

| Dataset | Path key | Used for |
|---------|----------|----------|
| **HydroRIVERS** | `paths$hydrorivers` | Directed `sfnetwork` edges; `hyriv_id`, `bb_id`, Strahler order |
| **FFR** | `paths$ffr_gdb` | `csi` and reach attributes joined to HydroRIVERS (`hyriv_id` = `reach_id`) |
| **GDW** | `paths$gdw_snapped` | Current dams (`is_current_dam = TRUE`); overlap filter vs FHReD |
| **FHReD** | `paths$fhred_snapped` | Future dams (`is_current_dam = FALSE`) |

# Load HydroRIVERS and FFR

```{r}
#| label: load-rivers-ffr
#| message: false
#| warning: false

rivers <- sf::read_sf(paths$hydrorivers) %>%
  janitor::clean_names() %>%
  sf::st_transform(target_crs)

ffr <- sf::read_sf(paths$ffr_gdb) %>%
  janitor::clean_names() %>%
  sf::st_transform(target_crs) %>%
  sf::st_cast("LINESTRING")
```

# Join HydroRIVERS + FFR

```{r}
#| label: join-rivers-ffr
#| message: false
#| warning: false

world_rivers_ffr <- rivers %>%
  dplyr::left_join(
    ffr %>% sf::st_drop_geometry(),
    by = c("hyriv_id" = "reach_id")
  )
```

# Load and prepare dam layers

```{r}
#| label: load-dams
#| message: false
#| warning: false

gdw <- sf::st_read(paths$gdw_snapped) %>% janitor::clean_names()

current_dams_gdw <- gdw %>%
  sf::st_make_valid() %>%
  sf::st_set_crs(target_crs) %>%
  sf::st_transform(target_crs) %>%
  dplyr::mutate(
    dam_id = as.character(.data$gdw_id),
    is_current_dam = TRUE
  )

future_dams_raw <- sf::st_read(paths$fhred_snapped) %>%
  sf::st_transform(target_crs) %>%
  dplyr::mutate(is_current_dam = FALSE) %>%
  as.data.frame() %>%
  janitor::clean_names()

future_dams_sf <- future_dams_raw %>%
  sf::st_as_sf(
    coords = c("lon_cleane", "lat_cleane"),
    crs = target_crs,
    remove = FALSE
  ) %>%
  sf::st_make_valid() %>%
  sf::st_transform(target_crs) %>%
  dplyr::mutate(
    dam_id = as.character(.data$dam_id),
    is_current_dam = FALSE
  )
```
