---
title: "net_with_dams_from_rivernetwork"
format: html
editor_options:
  chunk_output_type: inline
---

Builds a directed river network with the sfnetworks package using HydroRIVERS vectors, then blends GDW (current) + FHReD (future) dams to create the sfnetwork graph needed to calculate connectivity.

```{r}
#| label: setup
#| message: false
#| warning: false

library(sf)
library(sfnetworks)
library(tidygraph)
library(dplyr)
library(readr)
library(janitor)

target_crs <- sf::st_crs(3857) # We use EPSG 3857 for the entire analysis

# Load Global Dam Watch and prepare for blending
gdw <- st_read("/capstone/netzerohydro/data/GDW_prj_1000m_from_StrOrd3_SNAPPED.gpkg") %>% clean_names()

current_dams_gdw <- gdw %>%
  sf::st_make_valid() %>%
  sf::st_set_crs(target_crs) %>%
  sf::st_transform(target_crs) %>%
  dplyr::mutate(
    dam_id = as.character(.data$gdw_id),
    is_current_dam = TRUE
  )

# Confirm correct CRS
if (st_crs(current_dams_gdw)$epsg != target_crs) {
  message("Reprojecting to EPSG:3857")
  current_dams_gdw <- st_transform(current_dams_gdw, target_crs)
}

# Load Free-Flowing Rivers data
ffr <- read_sf("/capstone/netzerohydro/data/raw/FFR_river_network.gdb") %>% 
  st_transform(target_crs) %>% 
  clean_names()

# Load HydroRIVERS
rivers <- read_sf("/capstone/netzerohydro/data/raw/HydroRIVERS_v10_shp/HydroRIVERS_v10_shp") %>% 
  st_transform(target_crs) %>% 
  clean_names()

# Load HydroRIVERS and FFR
world_rivers_ffr <- rivers %>%
  left_join(
    ffr %>% # Raw ffr data
      st_drop_geometry(), # Remove ffr$geometry column (its the same as HydroRIVERS$Shape)
    by = c("hyriv_id" = "reach_id")
  ) %>% st_transform(target_crs)

# Confirm correct CRS
if (st_crs(world_rivers_ffr)$epsg != target_crs) {
    message(paste("Reprojecting to", target_crs))
    world_rivers_ffr <- st_transform(world_rivers_ffr, target_crs)
  }

# FHReD 
# Somewhere in out QGIS flow, some column names were messed up. This does not affect the workflow. 
future_dams_raw <- st_read("/capstone/netzerohydro/data/FhRED_prj_1000m_from_StrOrd3_SNAPPED.gpkg") |>
  st_transform(target_crs) %>%
  mutate(is_current_dam = FALSE) %>%
  as.data.frame() %>% 
  clean_names()

# Prepare FHReD future dams as sf points.
# st_as_sf() creates geometry from lon/lat columns.
future_dams_sf <- future_dams_raw %>%
  sf::st_as_sf(
    coords = c("lon_cleane", "lat_cleane"),
    crs = target_crs,
    remove = FALSE
  ) %>%
  sf::st_make_valid() %>%
  sf::st_transform(target_crs) %>%
  dplyr::mutate(
    dam_id = as.character(.data$dam_id),
    is_current_dam = FALSE
  )

# Confirm correct CRS before proceeding

if (st_crs(future_dams_sf)$epsg != target_crs) {
  message("Reprojecting to EPSG:3857")
  future_dams_sf <- st_transform(future_dams_sf, target_crs)
}

```

Define parameters for building river network. For this project we used the following parameters:
ord_stra_min <- 3
blend_tolerance_m <- Inf
fhred_overlap_tolerance_m <- 500
main_riv_filter <- NULL

```{r}
#| label: params
# -----------------------------
# Controls
# -----------------------------

# Keep only river reaches with Strahler order >.
ord_stra_min <- 2

# Snap tolerance in metres for st_network_blend().
blend_tolerance_m <- Inf 

# 1500mm is 1.5km - gives 3 pixel margin at 15-second arc gird

# Distance (metres) used to remove FHReD dams that are already represented
fhred_overlap_tolerance_m <- 500 # 6050 removed 205 dams

# Optional subset for testing:
# - NULL = all rivers
# - numeric/id value = keep only rows with this main_riv
main_riv_filter <- 70834639 # Mekong 41392598 # Amazon 60443230
```

Optional river filter to save compute time. Not needed for global analysis.

```{r}
#| label: filter-rivers
#| message: false
#| warning: false

# Filter HydroRIVERS by Strahler order
rivers_filtered <- world_rivers_ffr %>%
  dplyr::filter(.data$ord_stra > ord_stra_min)

# Optional: Filter by main_riv
if (!is.null(main_riv_filter)) {
  rivers_filtered <- rivers_filtered %>%
    dplyr::filter(as.numeric(.data$main_riv) == as.numeric(main_riv_filter))
}

# Ensure rivers are valid LINESTRING features for sfnetwork conversion.
rivers_filtered <- suppressWarnings(sf::st_cast(rivers_filtered, "LINESTRING"))
rivers_filtered <- rivers_filtered[!sf::st_is_empty(rivers_filtered), , drop = FALSE]

if (nrow(rivers_filtered) == 0L) {
  stop("No rivers left after filters. Check ord_stra/main_bas settings.", call. = FALSE)
}
```

```{r}
#| label: dedupe-fhred-vs-gdw
#| message: false
#| warning: false

# Keep only GDW rows explicitly labeled as FHReD-origin references.
gdw_fhred_ref <- current_dams_gdw %>%  dplyr::filter(.data$orig_src == "FHReD")

n_fhred_before <- nrow(future_dams_sf)

# st_is_within_distance() returns, for each FHReD row, indices of GDW ref points
# within the given tolerance. Any non-empty match means "already represented".
if (nrow(gdw_fhred_ref) > 0L && n_fhred_before > 0L) {
  overlap_idx <- sf::st_is_within_distance(
    future_dams_sf,
    gdw_fhred_ref,
    dist = fhred_overlap_tolerance_m
  )
  is_overlap <- lengths(overlap_idx) > 0L
  dams_removed <- sum(is_overlap)
  dams_future_fhred_clean <- future_dams_sf[!is_overlap, , drop = FALSE]
} else {
  dams_removed <- 0L
  dams_future_fhred_clean <- future_dams_sf
}


cat("FHReD total before de-duplication:", n_fhred_before, "\n")
cat("FHReD removed within", fhred_overlap_tolerance_m, "m of GDW orig_src='FHReD':", dams_removed, "\n")
cat("FHReD kept after de-duplication:", nrow(dams_future_fhred_clean), "\n")
```

FHReD and GDW have different primary keys with redundant IDs for dams. We add a prefix so the primary key in the combined dams can be the same, while retaining the same ID number. 

```{r}
#| label: Combine current & future dams
#| message: false
#| warning: false

# Too retain more attributes from FHReD and GDW add colnames to the transmute() below.

all_dams <- dplyr::bind_rows(
  dams_future_fhred_clean %>%
    dplyr::transmute(dam_id = paste0("FHRED_", dam_id), is_current_dam, geometry = sf::st_geometry(.), dam_name = project_na, capacity_mw = capacity, continent, country, main_basin = major_basi, main_river),
  current_dams_gdw %>%
    dplyr::transmute(dam_id = paste0("GDW_", dam_id), is_current_dam, geometry = sf::st_geometry(.), country, main_basin, dam_name = dam_name, capacity_mw = power_mw)
) %>% select(-geom) %>% 
  sf::st_as_sf(target_crs)
```

Build sfnetwork. 

```{r}
#| label: combine-and-build-network
#| message: false
#| warning: false

# In the current analysis we snapped in QGIS before blending. For the future, or reproducing, we would like to snap in R.

# all_dams_st <- st_snap(all_dams, rivers_filtered, tolerance = 1500) %>%
#   st_transform(target_crs)

# as_sfnetwork() converts river lines to a directed graph.
net <- sfnetworks::as_sfnetwork(rivers_filtered, directed = TRUE) %>%
  tidygraph::activate("edges") %>%
  # edge_length() computes geometric edge lengths; used as weighted distance input.
  dplyr::mutate(weight = sfnetworks::edge_length())


# st_network_blend() inserts/snap-aligns dam points onto nearest network edges.
net_with_dams <- sfnetworks::st_network_blend(net, all_dams, tolerance = blend_tolerance_m)
```

Now we need to pull the node and edge attributes from the sfnetwork dam network.

```{r}
#| label: snapped-dam-audit-with-river-attrs
#| message: false
#| warning: false

# Extract snapped network nodes and keep only dam nodes.
nodes_tbl <- net_with_dams %>%
  tidygraph::activate("nodes") %>%
  sf::st_as_sf() %>%
  dplyr::mutate(node_id = dplyr::row_number())

dams_snapped <- nodes_tbl %>%
  dplyr::filter(!is.na(.data$dam_id))

# Assign each snapped dam node to nearest filtered river reach.
# st_nearest_feature() returns row index of nearest river geometry per point.
nearest_riv_idx <- sf::st_nearest_feature(dams_snapped, rivers_filtered)

river_attr_keep <- c("hyriv_id", "bb_id", "ord_stra", "csi", "bas_name")
river_attr_keep <- river_attr_keep[river_attr_keep %in% names(rivers_filtered)]

dams_snapped <- dams_snapped %>%
  dplyr::bind_cols(
    rivers_filtered %>%
      sf::st_drop_geometry() %>%
      dplyr::slice(nearest_riv_idx) %>%
      dplyr::select(dplyr::all_of(river_attr_keep))
  )
```

For troubleshooting we save the all objects and check certain outputs. 

```{r}
#| label: diagnostics

n_current_total <- nrow(current_dams_gdw)
n_future_total <- nrow(dams_future_fhred_clean)
n_input_total <- nrow(all_dams)
n_snapped <- nrow(dams_snapped)
n_unsnapped <- n_input_total - n_snapped

cat("GDW current dams used:", n_current_total, "\n")
cat("FHReD future dams used (post-clean):", n_future_total, "\n")
cat("Total dams passed to blending:", n_input_total, "\n")
cat("Dams snapped onto network:", n_snapped, "\n")
cat("Dams not snapped:", n_unsnapped, "\n")
cat("Snapped future dams: ", sum(dams_snapped$is_current_dam %in% FALSE, na.rm = TRUE), "\n")
cat("Snapped current dams: ", sum(dams_snapped$is_current_dam %in% TRUE,  na.rm = TRUE), "\n")
```

Finally, save outputs as an object `out` to be ran with connectivity function. 

```{r}
#| label: output-objects

out <- list(
  net_with_dams = net_with_dams,
  rivers_filtered = rivers_filtered,
  dams_current_gdw = current_dams_gdw,
  dams_future_fhred_clean = dams_future_fhred_clean,
  all_dams = all_dams,
  dams_snapped = dams_snapped,
  main_riv_dist = main_riv_dist
)

file_name <- paste0("net_with_dams_GIS_v2.rds")
out_path <- file.path("/capstone/netzerohydro/data/processed_data", file_name)

saveRDS(out, out_path)
```

Optional build map to check out your network! WARNING: If global analysis was use, this will take a while. 

```{r}
library(leaflet)
library(sf)
library(dplyr)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolylines(
    data = out$net_with_dams %>%
      tidygraph::activate("edges") %>%
      sf::st_as_sf() %>%
      st_transform(4326),
    color = ~colorFactor("Set3", domain = bb_id)(as.character(bb_id)),
    weight = 2,
    opacity = 0.7,
    popup = ~paste0("bb_id: ", bb_id)
  ) %>%
  addCircleMarkers(
    data = out$net_with_dams %>%
      tidygraph::activate("nodes") %>%
      sf::st_as_sf() %>%
      dplyr::filter(!is.na(dam_id), is_current_dam %in% TRUE) %>%
      st_transform(4326),
    radius = 4, color = "blue", fillColor = "blue", fillOpacity = 0.9, stroke = FALSE,
    popup = ~paste0("Current dam<br>dam_id: ", dam_id)
  ) %>%
  addCircleMarkers(
    data = out$net_with_dams %>%
      tidygraph::activate("nodes") %>%
      sf::st_as_sf() %>%
      dplyr::filter(!is.na(dam_id), is_current_dam %in% FALSE) %>%
      st_transform(4326),
    radius = 4, color = "red", fillColor = "red", fillOpacity = 0.9, stroke = FALSE,
    popup = ~paste0("Future dam<br>dam_id: ", dam_id)
  ) %>%
  addScaleBar(position = "bottomleft")

```





