---
title: "1 — Build river network with blended dams"
format: html
editor_options:
  chunk_output_type: inline
---

## Prerequisites

Run `0_setup.qmd` in the same R session.

## Objects needed

`paths`, `target_crs`, `world_rivers_ffr`, `current_dams_gdw`, `future_dams_sf`

## Creates

`out` (optionally saved to `paths$net_with_dams_rds`)

Builds a directed river network with **sfnetworks** from HydroRIVERS + FFR reaches, blends GDW (current) and FHReD (future) dams, and saves `out` for connectivity analysis.

## Parameters

```{r}
#| label: params
# -----------------------------
# Controls
# -----------------------------

# Keep only river reaches with Strahler order >.
ord_stra_min <- 2

# Snap tolerance in metres for st_network_blend().
blend_tolerance_m <- Inf

# Distance (metres) used to remove FHReD dams already represented in GDW.
fhred_overlap_tolerance_m <- 500

# Optional subset for testing: NULL = all rivers; numeric = main_riv id only.
main_riv_filter <- 70834639 # Mekong 41392598 # Amazon 60443230
```

```{r}
#| label: filter-rivers
#| message: false
#| warning: false

rivers_filtered <- world_rivers_ffr %>%
  dplyr::filter(ord_stra > ord_stra_min)

if (!is.null(main_riv_filter)) {
  rivers_filtered <- rivers_filtered %>%
    dplyr::filter(as.numeric(main_riv) == as.numeric(main_riv_filter))
}

rivers_filtered <- suppressWarnings(sf::st_cast(rivers_filtered, "LINESTRING"))
rivers_filtered <- rivers_filtered[!sf::st_is_empty(rivers_filtered), , drop = FALSE]

if (nrow(rivers_filtered) == 0L) {
  stop("No rivers left after filters. Check ord_stra/main_riv settings.", call. = FALSE)
}
```

```{r}
#| label: dedupe-fhred-vs-gdw
#| message: false
#| warning: false

gdw_fhred_ref <- current_dams_gdw %>% dplyr::filter(orig_src == "FHReD")

n_fhred_before <- nrow(future_dams_sf)

if (nrow(gdw_fhred_ref) > 0L && n_fhred_before > 0L) {
  overlap_idx <- sf::st_is_within_distance(
    future_dams_sf,
    gdw_fhred_ref,
    dist = fhred_overlap_tolerance_m
  )
  is_overlap <- lengths(overlap_idx) > 0L
  dams_removed <- sum(is_overlap)
  dams_future_fhred_clean <- future_dams_sf[!is_overlap, , drop = FALSE]
} else {
  dams_removed <- 0L
  dams_future_fhred_clean <- future_dams_sf
}

cat("FHReD total before de-duplication:", n_fhred_before, "\n")
cat("FHReD removed within", fhred_overlap_tolerance_m, "m of GDW orig_src='FHReD':", dams_removed, "\n")
cat("FHReD kept after de-duplication:", nrow(dams_future_fhred_clean), "\n")
```

```{r}
#| label: combine-current-future-dams
#| message: false
#| warning: false

all_dams <- dplyr::bind_rows(
  dams_future_fhred_clean %>%
    dplyr::transmute(
      dam_id = paste0("FHRED_", dam_id),
      is_current_dam,
      geometry = sf::st_geometry(.),
      dam_name = project_na,
      capacity_mw = capacity,
      continent,
      country,
      main_basin = major_basi,
      main_river
    ),
  current_dams_gdw %>%
    dplyr::transmute(
      dam_id = paste0("GDW_", dam_id),
      is_current_dam,
      geometry = sf::st_geometry(.),
      country,
      main_basin,
      dam_name = dam_name,
      capacity_mw = power_mw
    )
) %>%
  dplyr::select(-dplyr::any_of("geom")) %>%
  sf::st_as_sf(crs = target_crs)
```

```{r}
#| label: combine-and-build-network
#| message: false
#| warning: false

net <- sfnetworks::as_sfnetwork(rivers_filtered, directed = TRUE) %>%
  tidygraph::activate("edges") %>%
  dplyr::mutate(weight = sfnetworks::edge_length())

net_with_dams <- sfnetworks::st_network_blend(net, all_dams, tolerance = blend_tolerance_m)
```

```{r}
#| label: snapped-dam-audit-with-river-attrs
#| message: false
#| warning: false

nodes_tbl <- net_with_dams %>%
  tidygraph::activate("nodes") %>%
  sf::st_as_sf() %>%
  dplyr::mutate(node_id = dplyr::row_number())

dams_snapped <- nodes_tbl %>%
  dplyr::filter(!is.na(dam_id))

nearest_riv_idx <- sf::st_nearest_feature(dams_snapped, rivers_filtered)

river_attr_keep <- c("hyriv_id", "bb_id", "ord_stra", "csi", "bas_name")
river_attr_keep <- river_attr_keep[river_attr_keep %in% names(rivers_filtered)]

dams_snapped <- dams_snapped %>%
  dplyr::bind_cols(
    rivers_filtered %>%
      sf::st_drop_geometry() %>%
      dplyr::slice(nearest_riv_idx) %>%
      dplyr::select(dplyr::all_of(river_attr_keep))
  )
```

```{r}
#| label: diagnostics

n_current_total <- nrow(current_dams_gdw)
n_future_total <- nrow(dams_future_fhred_clean)
n_input_total <- nrow(all_dams)
n_snapped <- nrow(dams_snapped)
n_unsnapped <- n_input_total - n_snapped

cat("GDW current dams used:", n_current_total, "\n")
cat("FHReD future dams used (post-clean):", n_future_total, "\n")
cat("Total dams passed to blending:", n_input_total, "\n")
cat("Dams snapped onto network:", n_snapped, "\n")
cat("Dams not snapped:", n_unsnapped, "\n")
cat("Snapped future dams: ", sum(dams_snapped$is_current_dam %in% FALSE, na.rm = TRUE), "\n")
cat("Snapped current dams: ", sum(dams_snapped$is_current_dam %in% TRUE, na.rm = TRUE), "\n")
```

```{r}
#| label: output-objects

out <- list(
  net_with_dams = net_with_dams,
  rivers_filtered = rivers_filtered,
  dams_current_gdw = current_dams_gdw,
  dams_future_fhred_clean = dams_future_fhred_clean,
  all_dams = all_dams,
  dams_snapped = dams_snapped
)

saveRDS(out, paths$net_with_dams_rds)
message("Saved network output to: ", paths$net_with_dams_rds)
```

Optional map (slow for global networks):

```{r}
#| label: optional-map
#| eval: false

leaflet::leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  leaflet::addPolylines(
    data = out$net_with_dams %>%
      tidygraph::activate("edges") %>%
      sf::st_as_sf() %>%
      sf::st_transform(4326),
    color = ~ leaflet::colorFactor("Set3", domain = bb_id)(as.character(bb_id)),
    weight = 2,
    opacity = 0.7,
    popup = ~ paste0("bb_id: ", bb_id)
  ) %>%
  leaflet::addCircleMarkers(
    data = out$net_with_dams %>%
      tidygraph::activate("nodes") %>%
      sf::st_as_sf() %>%
      dplyr::filter(!is.na(dam_id), is_current_dam %in% TRUE) %>%
      sf::st_transform(4326),
    radius = 4, color = "blue", fillColor = "blue", fillOpacity = 0.9, stroke = FALSE,
    popup = ~ paste0("Current dam<br>dam_id: ", dam_id)
  ) %>%
  leaflet::addCircleMarkers(
    data = out$net_with_dams %>%
      tidygraph::activate("nodes") %>%
      sf::st_as_sf() %>%
      dplyr::filter(!is.na(dam_id), is_current_dam %in% FALSE) %>%
      sf::st_transform(4326),
    radius = 4, color = "red", fillColor = "red", fillOpacity = 0.9, stroke = FALSE,
    popup = ~ paste0("Future dam<br>dam_id: ", dam_id)
  ) %>%
  leaflet::addScaleBar(position = "bottomleft")
```
