---
title: "Using connectivity on a different dataset"
format: html
---

## Idea

`connectivity_from_network()` does not read files. It only needs a ready-made **`net_with_dams`**: a directed `sfnetwork` with dams snapped onto it.

## What your network must have

**Edges**

- `weight` — edge length in **metres** (used as km in the function)
- `bb_id` — trunk / reach ID for hop-based tie-breaking at confluences

**Nodes (dam points only)**

- `dam_id` — unique ID (character)
- `is_current_dam` — `TRUE` for existing dams, `FALSE` for future/planned dams

## Minimal example

```{r}
#| eval: false

library(sf)
library(sfnetworks)
library(tidygraph)
library(dplyr)

# 1. Your river lines (LINESTRING, one CRS for everything)
# rivers_sf needs a geometry column and bb_id per reach

net <- sfnetworks::as_sfnetwork(rivers_sf, directed = TRUE) %>%
  tidygraph::activate("edges") %>%
  dplyr::mutate(weight = sfnetworks::edge_length())

# 2. Your dam points — must include dam_id and is_current_dam
# dams_sf <- sf object with POINT geometry

net_with_dams <- sfnetworks::st_network_blend(net, dams_sf, tolerance = Inf)

# 3. Run connectivity
source("R/connectivity_from_network.R")

out_conn <- connectivity_from_network(
  net_with_dams = net_with_dams,
  threshold_downstream_km = 250,
  threshold_upstream_km = 250,
  threshold_cascade_km = 250
)

reach_df <- out_conn$reach_df
```

## Checklist

1. Project rivers and dams to the **same CRS**.
2. Ensure edge **`weight`** is in metres and **`bb_id`** exists on every edge used in the analysis.
3. Tag each dam node with **`dam_id`** and **`is_current_dam`** before blending.
4. Use `st_network_blend()` (or equivalent snapping) so dam points sit on network nodes.
5. Pick km thresholds that match your study extent.

## Optional enrichment

After connectivity, append river attributes if your edges carry them:

```{r}
#| eval: false

source("R/add_ffr_attr.R")

reach_enriched <- add_ffr_attr(
  net_with_dams = net_with_dams,
  reach_df = out_conn$reach_df,
  edge_attr_cols = c("csi", "bas_name")  # use column names from your edges
)
```

## Full pipeline reference

For the GDW + FHReD + HydroRIVERS workflow, follow notebooks `0_setup.qmd` through `3_enrich_out_conn.qmd`.
