---
title: "Function testing: basin network + connectivity (cascade_level)"
format: html
editor_options:
  chunk_output_type: inline
---

This document is an **end-to-end test + debug viewer** for:

- `net_with_dams_from_hydrobasin()` (build a basin network + snap dams)
- `connectivity_from_network()` (returns `list(reach_df=..., debug=...)`)
- `add_ffr_attr()` (joins river-edge attributes onto `reach_df` by `bb_id`)

It also prints the key intermediate objects created inside `connectivity_from_network()` so you can inspect them in the rendered HTML:

- `ds_dam`: downstream **current** dams from each future dam (future → current)
- `us_dam`: upstream **current** dams from each future dam (current → future)
- `node_bb`: node → `bb_id` (trunk) mapping
- `trunk_edges`: inferred trunk adjacency edges (confluence trunk graph)

---

```{r}
#| label: setup
#| output: false

# Packages: spatial (sf), networks (sfnetworks, tidygraph, igraph), tables (dplyr),
# plotting (ggplot2, ggnewscale), data IO (readr), name cleaning (janitor).
library(sfnetworks)
library(sf)
library(tidygraph)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(igraph)
library(ggnewscale)

# Load the two scripts so their functions exist in this Quarto session.
source("connectivity_function_v2.R")
source("net_with_dams_from_hydrobasin.R")
source("add_ffr_attr.R")
```

```{r}
#| label: load-data
#| message: false
#| warning: false

# Current dams (GRanD).
current_dams <- read_sf("../../../../../capstone/netzerohydro/data/raw/GRanD_Version_1_3") |>
  clean_names()

# Future dams (FHReD table; must include lon_cleaned/lat_cleaned).
future_dams <- read_csv("../../../../../capstone/netzerohydro/data/raw/FHReD_2015_future_dams/FHReD_2015_future_dams_Zarfl_et_al_beta_version.csv") |>
  clean_names()

# HydroBASINS polygons (level must match your hybas_id).
basins_sf <- read_sf("../../../../../capstone/netzerohydro/data/raw/hybas_as_lev01-12_v1c/hybas_as_lev04_v1c.shp") |>
  clean_names()

# Global river + FFR attributes.
world_rivers_ffr <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/world_rivers_ffr.rds")
```

```{r}
#| label: build-network-and-connectivity
#| message: false
#| warning: false

# Change selected basin here.
hybas_id <- 4041043580
continent <- "Asia"

# Build study network (one Hybas id + crop + optional blending).
out <- net_with_dams_from_hydrobasin(
  current_dams      = current_dams,
  future_dams       = future_dams,
  basins_sf         = basins_sf,
  world_rivers_ffr  = world_rivers_ffr,
  hybas_id          = hybas_id,
  continent         = continent,
  ord_stra_min      = 3L,
  blend_tolerance_m = 50000
)

# OR read exisiting net_with_dams output from workbench
#out <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/nets_with_dams/net_hybas4020015090_Asia_pfaf02.rds") |>
 # clean_names()

# Run connectivity (returns list with reach_df + debug).
out_conn <- connectivity_from_network(
  out$net_with_dams,
  cascade_threshold_km = 100
)

# Extract the reach table (standalone data frame).
reach_df <- out_conn$reach_df

# Extract debug objects (intermediates you can inspect).
dbg <- out_conn$debug

# Quick preview.
head(reach_df)
```

```{r}
#| label: build-dataframe-with-ffr-attributes
#| message: false
#| warning: false

# Join river-edge attributes by bb_id.
reach_enriched <- add_ffr_attr(
  out$net_with_dams,
  reach_df,
  edge_attr_cols = c("csi", "ord_stra", "bas_name"),
  future_extra   = future_dams %>%
  dplyr::mutate(dam_id = as.character(.data$dam_id)) %>%
  dplyr::select(dam_id, capacity_mw, project_name)
)
```

```{r}
#| label: debug-step-1-dams

# Step 1: dam nodes (what the network thinks are dams).
head(dbg$dam_nodes_tbl, 20)

# Current dam nodes.
head(dbg$current_dams_tbl, 20)

# Future dam nodes.
head(dbg$future_dams_tbl, 20)
```

```{r}
#| label: debug-step-2-node-bb

# Step 2: node -> bb_id (trunk) mapping.
head(dbg$node_bb, 20)
```

```{r}
#| label: debug-step-3-all-pairs

# Step 3: ALL future/current dam pairs with distances.
# Downstream = future -> current.
head(dbg$ds_dam_all, 20)

# Upstream = current -> future (still stored as future/current pairs).
head(dbg$us_dam_all, 20)
```

```{r}
#| label: debug-step-4-downstream-filter

# Step 4: downstream filtering (strict same trunk).
head(dbg$ds_reachable, 20)
head(dbg$ds_same_trunk, 20)

# Per-future downstream summary.
head(dbg$downstream_summary, 20)

# Chosen downstream dam per future (closest same-trunk).
head(dbg$chosen_downstream, 20)
```

```{r}
#| label: debug-step-5-trunk-graph

# Step 5: trunk adjacency graph inferred at confluences.
head(dbg$trunk_edges, 50)
dbg$g_trunk
```

```{r}
#| label: debug-step-6-upstream-filter

# Step 6: upstream filtering within threshold + trunk-step distances.
head(dbg$us_reachable, 20)
head(dbg$us_within_threshold, 20)
head(dbg$us_with_steps, 20)

# Chosen upstream dam per future (closest within threshold).
head(dbg$chosen_upstream, 20)
```

```{r}
#| label: debug-step-7-final-output

# Step 7: final reach_df (what you export / join later).
head(reach_df, 20)
```

```{r}
#| label: debug-step-8-decision-table

# Step 8: one-row-per-future decision table (chosen up/downstream + cascade_level).
head(dbg$decision_table, 50)
```

```{r}
#| label: debug-ds-us

# Look at the downstream dam table (future -> current).
head(dbg$ds_dam_all, 20)

# Look at the upstream dam table (current -> future).
head(dbg$us_dam_all, 20)
```

```{r}
#| label: debug-bb

# Node -> bb_id mapping (trunk assignment).
head(dbg$node_bb, 20)

# Future dam -> bb_id.
head(dbg$future_trunks, 20)

# Current dam -> bb_id.
head(dbg$current_trunks, 20)
```

```{r}
#| label: debug-trunk-graph

# Look at trunk adjacency edges inferred at confluences.
head(dbg$trunk_edges, 50)

# Trunk graph summary.
dbg$g_trunk
```

```{r}
#| label: map
#| fig-width: 10
#| fig-height: 7

# Interactive map (mapview). Requires `out` and `reach_df` from earlier chunks.

library(mapview)
library(sf)
library(dplyr)

# Basin polygon.
one_subbasin <- out$basin_poly

# Cropped rivers.
rivers_one <- out$rivers_cropped

# IMPORTANT: use the SNAPPED/BLENDED dam nodes from the network so the map
# shows the exact same dam set as `reach_df`.
dam_nodes_sf <- out$net_with_dams %>%
  tidygraph::activate("nodes") %>% # switch to nodes
  sf::st_as_sf() %>% # node points as sf
  dplyr::filter(!is.na(.data$dam_id)) %>% # keep only dam nodes
  dplyr::mutate(dam_id = as.character(.data$dam_id)) # ensure join key is character

# Split dam nodes into future vs current using the node flag.
dams_fut_nodes <- dplyr::filter(dam_nodes_sf, .data$is_current_dam %in% FALSE)
dams_cur_nodes <- dplyr::filter(dam_nodes_sf, .data$is_current_dam %in% TRUE)

# Keep only future dams from reach_df (cascade_level is defined for future dams).
reach_future <- dplyr::filter(reach_df, .data$dam_type == "future")

# Join cascade_level from reach_df onto the snapped future dam nodes.
dams_fut <- dplyr::left_join(
  dams_fut_nodes,
  dplyr::select(reach_future, dam_id, cascade_level),
  by = "dam_id"
)

# Cascade definition for mapping: cascade_level >= 1.
dams_future_cascade <- dplyr::filter(
  dams_fut,
  !is.na(.data$cascade_level) & .data$cascade_level >= 1L
)

# Future dams that are not in cascade.
dams_future_other <- dplyr::filter(
  dams_fut,
  is.na(.data$cascade_level) | .data$cascade_level < 1L
)

# Current dams from snapped nodes (if any).
dams_cur <- if (nrow(dams_cur_nodes) > 0L) dams_cur_nodes else NULL

# Hover labels for rivers.
riv_df <- sf::st_drop_geometry(rivers_one)
lab_rivers <- paste0("bb_id: ", riv_df$bb_id)

if ("ord_stra" %in% names(riv_df)) lab_rivers <- paste0(lab_rivers, " | ord_stra: ", riv_df$ord_stra)
if ("csi" %in% names(riv_df)) lab_rivers <- paste0(lab_rivers, " | csi: ", riv_df$csi)
if ("bas_name" %in% names(riv_df)) lab_rivers <- paste0(lab_rivers, " | bas_name: ", riv_df$bas_name)

# Base layers: basin and rivers.
mv <- mapview(
  one_subbasin,
  color = "black",
  alpha.regions = 0,
  legend = FALSE,
  layer.name = "Basin"
) +
  mapview(
    rivers_one,
    zcol = "bb_id",
    legend = FALSE,
    label = lab_rivers,
    layer.name = "Rivers"
  )

# Current dams layer.
if (!is.null(dams_cur) && nrow(dams_cur) > 0L) {
  dcur <- sf::st_drop_geometry(dams_cur)
  lab_cur <- paste0("dam_id: ", dcur$dam_id, " | current")

  mv <- mv + mapview(
    dams_cur,
    col.regions = "red",
    color = "red",
    legend = FALSE,
    label = lab_cur,
    layer.name = "Current dams"
  )
}

# Future non-cascade layer.
if (nrow(dams_future_other) > 0L) {
  oth <- sf::st_drop_geometry(dams_future_other)
  lab_other <- paste0("dam_id: ", oth$dam_id, " | cascade_level: ", oth$cascade_level)

  mv <- mv + mapview(
    dams_future_other,
    col.regions = "gray60",
    color = "gray60",
    legend = FALSE,
    label = lab_other,
    layer.name = "Future (not cascade)"
  )
}

# Future cascade layer (colored by cascade_level).
if (nrow(dams_future_cascade) > 0L) {
  cas <- sf::st_drop_geometry(dams_future_cascade)
  lab_cas <- paste0("dam_id: ", cas$dam_id, " | cascade_level: ", cas$cascade_level)

  mv <- mv + mapview(
    dams_future_cascade,
    zcol = "cascade_level",
    legend = TRUE,
    label = lab_cas,
    layer.name = "Future (cascade)"
  )
}

mv
```



```{r}
#| label: leaflet-from-reach-enriched
#| message: false
#| warning: false

library(leaflet)
library(sf)
library(dplyr)
library(htmltools)

if (!exists("out", inherits = FALSE)) stop("Missing object: out", call. = FALSE)
if (!exists("reach_enriched", inherits = FALSE)) stop("Missing object: reach_enriched", call. = FALSE)

# Prepare attribute table
re <- reach_enriched %>%
  mutate(
    dam_id = as.character(dam_id),
    dam_type = as.character(dam_type),
    cascade_level = as.integer(cascade_level)
  )

# Split future/current attributes
re_future <- re %>% filter(dam_type == "future")
re_current <- re %>% filter(dam_type == "current")

# Spatial layers (WGS84)
basin_wgs  <- st_transform(out$basin_poly, 4326)
rivers_wgs <- st_transform(out$rivers_cropped, 4326)

current_wgs <- out$dams_current_cropped %>%
  mutate(dam_id = as.character(dam_id)) %>%
  left_join(re_current, by = "dam_id", suffix = c("", ".re")) %>%
  st_transform(4326)

future_wgs <- out$dams_future_cropped %>%
  mutate(dam_id = as.character(dam_id)) %>%
  inner_join(re_future, by = "dam_id", suffix = c("", ".re")) %>%
  st_transform(4326)

# ---- UTF-8 + HTML safety for leaflet payload ----
fix_utf8 <- function(x) {
  if (!is.character(x)) return(x)
  y <- iconv(x, from = "", to = "UTF-8", sub = "")
  y[is.na(y)] <- ""
  y
}

current_wgs <- current_wgs %>% mutate(across(where(is.character), fix_utf8))
future_wgs  <- future_wgs  %>% mutate(across(where(is.character), fix_utf8))

esc <- function(x) htmltools::htmlEscape(as.character(x))

# Color future by cascade level
pal <- colorFactor(
  palette = c(
    "0" = "#9E9E9E",  # no cascade
    "1" = "#A0FCA0",
    "2" = "#63BD5D",
    "3" = "#255915"
  ),
  domain = as.character(sort(unique(future_wgs$cascade_level))),
  na.color = "#BDBDBD"
)

future_popup <- sprintf(
  paste0(
    "<b>Future dam</b><br/>dam_id: %s<br/>project: %s<br/>",
    "cascade_level: %s<br/>us_trunks_away: %s<br/>",
    "has_upstream: %s<br/>has_downstream: %s<br/>",
    "min_up_km: %s<br/>min_down_km: %s<br/>csi: %s"
  ),
  esc(future_wgs$dam_id),
  esc(if ("project_name" %in% names(future_wgs)) future_wgs$project_name else NA_character_),
  esc(future_wgs$cascade_level),
  esc(future_wgs$us_trunks_away),
  esc(future_wgs$has_current_upstream),
  esc(future_wgs$has_current_downstream),
  esc(round(future_wgs$min_distance_upstream_km, 2)),
  esc(round(future_wgs$min_distance_downstream_km, 2)),
  esc(round(future_wgs$csi, 2))
)

current_popup <- sprintf(
  "<b>Current dam</b><br/>dam_id: %s<br/>name: %s",
  esc(current_wgs$dam_id),
  esc(if ("dam_name" %in% names(current_wgs)) current_wgs$dam_name else NA_character_)
)

current_popup <- sprintf(
  "<b>Current dam</b><br/>dam_id: %s<br/>name: %s",
  current_wgs$dam_id,
  if ("dam_name" %in% names(current_wgs)) current_wgs$dam_name else NA_character_
)

leaflet(options = leafletOptions(preferCanvas = TRUE)) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    data = basin_wgs,
    color = "#222222", weight = 1.5, fill = FALSE,
    group = "Basin"
  ) %>%
  addPolylines(
    data = rivers_wgs,
    color = "#4C78A8", weight = 1, opacity = 0.5,
    group = "Rivers"
  ) %>%
  addCircleMarkers(
    data = current_wgs,
    radius = 3.5, weight = 1,
    color = "red", fillColor = "red", fillOpacity = 0.85,
    popup = current_popup,
    group = "Current dams"
  ) %>%
  addCircleMarkers(
    data = future_wgs,
    radius = 5, weight = 1,
    color = ~pal(as.character(cascade_level)),
    fillColor = ~pal(as.character(cascade_level)),
    fillOpacity = 0.95,
    popup = future_popup,
    group = "Future dams"
  ) %>%
  addLegend(
    "bottomright",
    pal = pal,
    values = as.character(future_wgs$cascade_level),
    title = "Future dam cascade_level",
    opacity = 1
  ) %>%
  addLayersControl(
    overlayGroups = c("Basin", "Rivers", "Current dams", "Future dams"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addScaleBar(position = "bottomleft") %>%
  fitBounds(
    lng1 = st_bbox(basin_wgs)[["xmin"]],
    lat1 = st_bbox(basin_wgs)[["ymin"]],
    lng2 = st_bbox(basin_wgs)[["xmax"]],
    lat2 = st_bbox(basin_wgs)[["ymax"]]
  )
```

