---
title: "Basin network + connectivity (demo)"
format: html
editor_options:
  chunk_output_type: inline
---

1. `source()` connectivity_function.R and basin_network.R
2. `net_with_dams_from_hydrobasin()` — **`continent`** (must match `world_rivers_ffr$continent`) plus **Strahler**, then **crop to the basin** so `st_filter` does not scan the whole global layer. **`future_dams`** is **`data.frame`** with **`lon_cleaned`** / **`lat_cleaned`**.
3. `connectivity_from_network()` then `add_ffr_attr()` (`edge_attr_cols` = add FFR columns from river edges by `bb_id`; `future_extra` optional; `output_cols` = keep only listed columns). Map at the end.

---

```{r}
#| label: setup
#| output: false

# Packages: spatial (sf), networks (sfnetworks, tidygraph, igraph), tables (dplyr),
# plotting (ggplot2, ggnewscale), data IO (readr), name cleaning (janitor).
library(sfnetworks)
library(sf)
library(tidygraph)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(igraph)
library(ggnewscale)

# Load the two scripts
source("connectivity_function.R")
source("basin_network.R")
```

```{r}
#| label: load-data
#| message: false
#| warning: false

current_dams <- read_sf("../../../../../capstone/netzerohydro/data/raw/GRanD_Version_1_3") |>
  clean_names()

# lon/lat columns expected by net_with_dams_from_hydrobasin).
future_dams <- read_csv("../../../../../capstone/netzerohydro/data/raw/FHReD_2015_future_dams/FHReD_2015_future_dams_Zarfl_et_al_beta_version.csv") |>
  clean_names()

# Global river + FFR attributes
world_rivers_ffr <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/world_rivers_ffr.rds")


```

```{r}
#| label: sub-basin-exploration

# Hybas polygons to use
basins_sf <- read_sf("../../../../../capstone/netzerohydro/data/raw/hybas_af_lev04_v1c/hybas_af_lev04_v1c.shp") |>
  clean_names()

# Explore basins in the loaded file
basins_sf_big <- basins_sf %>% filter(basins_sf$sub_area > 100000)
head(basins_sf_big, 14)

ggplot() +
  geom_sf(data = basins_sf_big, fill = NA, color = "black", linewidth = 0.4) +
  geom_sf(
    data = basins_sf_big,
    aes(),
    linewidth = 0.35,
    alpha = 0.9
  ) +
  facet_wrap(~ hybas_id) +   # drop scales = "free"
  guides(color = "none") +
  theme_minimal() +
  labs(
    title = "sub-basins",
    subtitle = "One panel per sub-basin"
  )
```

```{r}
#| label: Interactive Map 

library(tidyverse)
library(plotly)
library(htmlwidgets)
library(scales)
library(dplyr)

ggplot_static <-basins_sf %>% 
  #filter(hybas_id == 1040000010) %>% 
  ggplot() + 
  geom_sf(aes(text = paste0(
            "hybas_id: ", hybas_id))) + 
  theme_void()

# interactive plot 
ggplotly(ggplot_static, 
         tooltip = "text" # inside of hover tool tip
         )

```

```{r}
#| label: Get basin's hybas_id as a csv  

# Save your basin as a csv 
#write_csv(basins_sf, here::here("africa_basin_sf.csv"))

# Read the csv in 
#basin_list <- read_csv(here::here("africa_basin_sf.csv"))

```


```{r}
#| label: Looping thru function and saving data  

# For loop 
for (hybas_id in basin_list$hybas_id) {
 
  continent <- "Africa"
  
  out <- net_with_dams_from_hydrobasin(
    current_dams      = current_dams,
    future_dams       = future_dams,
    basins_sf         = basins_sf,
    world_rivers_ffr  = world_rivers_ffr,
    hybas_id          = hybas_id,
    continent         = continent,
    ord_stra_min      = 3L,
    blend_tolerance_m = 50000)
    
    if(is.null(out)){
      message("Skipping:",hybas_id)
      next
    }
  
    saveRDS(
      out,
      sprintf(
        "../../../../../capstone/netzerohydro/data/cleaned_v1/nets_with_dams/net_hybas%s_%s_pfaf%s.rds",
        hybas_id,
        continent,
        substr(as.character(hybas_id), 2, 3)
      )
    )
}
```

```{r}
#| label: build-network-and-connectivity

# Change Selected Basin Here!!!
hybas_id <- 1040003990
continent <- "Africa" # From list below

# > unique(world_rivers_ffr$continent)
# "Africa"        "Europe"        "Asia"          "Australia"    
# "South America" "North America"

# HydroBASINS Regions Artic, Greenland and Siberia (might be Asia) are not in FFR data (This may need to be addressed later... hopefully not)

# Build study network: one Hybas id, continent matches river layer, blend tolerance in metres.
out <- net_with_dams_from_hydrobasin(
  current_dams      = current_dams,
  future_dams       = future_dams,
  basins_sf         = basins_sf,
  world_rivers_ffr  = world_rivers_ffr,
  hybas_id          = hybas_id,
  continent         = continent,
  ord_stra_min      = 3L,
  blend_tolerance_m = 50000
)

# Save network as an RDS file to not have to run over the same RDS each time
saveRDS(out, sprintf("../../../../../capstone/netzerohydro/data/cleaned_v1/nets_with_dams/net_hybas%s_%s_pfaf%s.rds", hybas_id, continent, substr(as.character(hybas_id), 2, 3)))

# OPTIONAL READ RDS (out is for output)
#out <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/nets_with_dams/net_hybas4040960480_Asia_pfaf04.rds")
```


```{r}
#| label: build-dataframe-with-ffr-attributes


# Per-dam connectivity and cascade labels; trunk_steps = 0 means same bb_id only.
reach_df <- connectivity_from_network(
  out$net_with_dams,
  cascade_threshold_km = NULL,
  cascade_trunk_steps  = 5L
)

# --- reach_enriched ---
# edge_attr_cols: pull these columns FROM RIVER EDGES (join on bb_id).
# future_extra: optional dam-level table by dam_id; omit (NULL)
# output_cols: only SELECT these names at the end (REACH_DF_OUTPUT_COLS = connectivity core).
reach_output_cols <- c(
  REACH_DF_OUTPUT_COLS,
  "csi", "ord_stra", "bas_name"
)

reach_enriched <- add_ffr_attr(
  out$net_with_dams,
  reach_df,
  edge_attr_cols = c("csi", "ord_stra", "bas_name"),
  future_extra   = NULL,
  output_cols    = reach_output_cols
)

head(reach_enriched)
```

```{r}
#| label: map
#| fig-width: 10
#| fig-height: 7

# Interactive map (mapview). Requires `out` and `reach_df` from earlier chunks.

library(mapview)
library(sf)

one_subbasin <- out$basin_poly
rivers_one <- out$rivers_cropped

reach_future <- dplyr::filter(reach_df, .data$dam_type == "future")

dams_fut <- dplyr::left_join(
  out$dams_future_cropped,
  dplyr::select(reach_future, dam_id, cascade_status),
  by = "dam_id"
)

dams_future_cascade <- dplyr::filter(dams_fut, .data$cascade_status == "cascade")
dams_future_other <- dplyr::filter(
  dams_fut,
  is.na(.data$cascade_status) | .data$cascade_status != "cascade"
)

dams_cur <- if (nrow(out$dams_current_cropped) > 0L) {
  sf::st_point_on_surface(out$dams_current_cropped)
} else {
  NULL
}

# Hover text: explicit string
riv_df <- sf::st_drop_geometry(rivers_one)
lab_rivers <- paste0("bb_id: ", riv_df$bb_id)

if ("ord_stra" %in% names(riv_df)) {
  lab_rivers <- paste0(lab_rivers, " | ord_stra: ", riv_df$ord_stra)
}
if ("csi" %in% names(riv_df)) {
  lab_rivers <- paste0(lab_rivers, " | csi: ", riv_df$csi)
}
if ("bas_name" %in% names(riv_df)) {
  lab_rivers <- paste0(lab_rivers, " | bas_name: ", riv_df$bas_name)
}

mv <- mapview(
  one_subbasin,
  color = "black",
  alpha.regions = 0,
  legend = FALSE,
  layer.name = "Basin"
) +
  mapview(
    rivers_one,
    zcol = "bb_id",
    legend = FALSE,
    label = lab_rivers,
    layer.name = "Rivers"
  )

if (!is.null(dams_cur) && nrow(dams_cur) > 0L) {
  dcur <- sf::st_drop_geometry(dams_cur)
  lab_cur <- paste0("dam_id: ", dcur$dam_id, " | current")
  if ("grand_id" %in% names(dcur)) {
    lab_cur <- paste0(lab_cur, " | grand_id: ", dcur$grand_id)
  }
  mv <- mv + mapview(
    dams_cur,
    col.regions = "red",
    color = "red",
    legend = FALSE,
    label = lab_cur,
    layer.name = "Current dams"
  )
}

if (nrow(dams_future_other) > 0L) {
  oth <- sf::st_drop_geometry(dams_future_other)
  lab_other <- paste0("dam_id: ", oth$dam_id, " | cascade: ", oth$cascade_status)
  mv <- mv + mapview(
    dams_future_other,
    col.regions = "gray60",
    color = "gray60",
    legend = FALSE,
    label = lab_other,
    layer.name = "Future (not cascade)"
  )
}

if (nrow(dams_future_cascade) > 0L) {
  cas <- sf::st_drop_geometry(dams_future_cascade)
  lab_cas <- paste0("dam_id: ", cas$dam_id, " | cascade: ", cas$cascade_status)
  mv <- mv + mapview(
    dams_future_cascade,
    zcol = "dam_id",
    legend = FALSE,
    label = lab_cas,
    layer.name = "Future (cascade)"
  )
}

mv
```
