---
title: "Basin network + connectivity (demo)"
format: html
editor_options:
  chunk_output_type: inline
---

1. `source()` connectivity_function.R and basin_network.R
2. `net_with_dams_from_hydrobasin()` — **`continent`** (must match `world_rivers_ffr$continent`) plus **Strahler**, then **crop to the basin** so `st_filter` does not scan the whole global layer. **`future_dams`** is **`data.frame`** with **`lon_cleaned`** / **`lat_cleaned`**.
3. `connectivity_from_network()` then `add_ffr_attr()` (`edge_attr_cols` = add FFR columns from river edges by `bb_id`; `future_extra` optional; `output_cols` = keep only listed columns). Map at the end.

---

```{r}
#| label: setup
#| output: false

# Packages: spatial (sf), networks (sfnetworks, tidygraph, igraph), tables (dplyr),
# plotting (ggplot2, ggnewscale), data IO (readr), name cleaning (janitor).
library(sfnetworks)
library(sf)
library(tidygraph)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(igraph)
library(ggnewscale)

# Load the two scripts so their functions exist in this Quarto session.
source("connectivity_function.R")
source("basin_network.R")
```

```{r}
#| label: load-data

# Paths point to your project’s raw/cleaned data; adjust if your folder layout differs.
current_dams <- read_sf("../../../../../capstone/netzerohydro/data/raw/GRanD_Version_1_3") |>
  clean_names()

# Table of future dam sites (lon/lat columns expected by net_with_dams_from_hydrobasin).
future_dams <- read_csv("../../../../../capstone/netzerohydro/data/raw/FHReD_2015_future_dams/FHReD_2015_future_dams_Zarfl_et_al_beta_version.csv") |>
  clean_names()

# Hybas polygons for South America; level (e.g. lev04) must match hybas_id.
basins_sf <- read_sf("../../../../../capstone/netzerohydro/data/raw/hybas_sa_lev01-12_v1c/hybas_sa_lev04_v1c.shp") |>
  clean_names()

# Global river + FFR attributes; must include continent, ord_stra, bb_id for the pipeline.
world_rivers_ffr <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/world_rivers_ffr.rds")
```

```{r}
#| label: build-network-and-connectivity

# Build study network: one Hybas id, continent matches river layer, blend tolerance in metres.
out <- net_with_dams_from_hydrobasin(
  current_dams      = current_dams,
  future_dams       = future_dams,
  basins_sf         = basins_sf,
  world_rivers_ffr  = world_rivers_ffr,
  hybas_id          = 6040345000,
  continent         = "South America",
  ord_stra_min      = 3L,
  blend_tolerance_m = 50000
)

# Per-dam connectivity and cascade labels; trunk_steps = 0 means same bb_id only.
reach_df <- connectivity_from_network(
  out$net_with_dams,
  cascade_threshold_km = 500,
  cascade_trunk_steps  = 0L
)

# --- reach_enriched ---
# edge_attr_cols: pull these columns FROM RIVER EDGES (join on bb_id).
# future_extra: optional dam-level table by dam_id; omit (NULL)
# output_cols: only SELECT these names at the end (REACH_DF_OUTPUT_COLS = connectivity core).
reach_output_cols <- c(
  REACH_DF_OUTPUT_COLS,
  "csi", "ord_stra", "bas_name"
)

reach_enriched <- add_ffr_attr(
  out$net_with_dams,
  reach_df,
  edge_attr_cols = c("csi", "ord_stra", "bas_name"),
  future_extra   = NULL,
  output_cols    = reach_output_cols
)

head(reach_enriched)
```

```{r}
#| label: map
#| fig-width: 10
#| fig-height: 7

# Study basin outline (from HydroBASINS); hollow fill so rivers show through.
one_subbasin <- out$basin_poly

# Rivers colored by coarse trunk id: first 4 chars of bb_id group segments for legibility.
rivers_one <- out$rivers_cropped %>%
  dplyr::mutate(bb_id_prefix = substr(as.character(.data$bb_id), 1, 4))

# Current dams as points (point-on-surface avoids odd multipoint geometries from intersect).
dams_current_one <- if (nrow(out$dams_current_cropped) > 0L) {
  sf::st_point_on_surface(out$dams_current_cropped)
} else {
  out$dams_current_cropped
}

# connectivity_from_network() appends current dams with NA cascade fields — map only future for cascade join.
reach_future <- dplyr::filter(reach_df, .data$dam_type == "future")

# Attach cascade label to cropped future dam sf for splitting plot layers.
dams_fut <- out$dams_future_cropped %>%
  dplyr::left_join(
    dplyr::select(reach_future, dam_id, cascade_status),
    by = "dam_id"
  )

# Split future dams: cascade vs not (for color emphasis on cascade only).
dams_future_cascade <- dams_fut %>%
  dplyr::filter(.data$cascade_status == "cascade")

dams_future_other <- dams_fut %>%
  dplyr::filter(is.na(.data$cascade_status) | .data$cascade_status != "cascade")

# One color per cascade dam id (only build palette if any exist).
cascade_ids <- unique(as.character(dams_future_cascade$dam_id))
if (length(cascade_ids) > 0L) {
  cascade_pal <- stats::setNames(
    scales::hue_pal()(length(cascade_ids)),
    cascade_ids
  )
} else {
  cascade_pal <- character(0)
}

# Base plot: basin boundary, rivers (first color scale = bb_id prefix), current + non-cascade future in fixed colors.
p <- ggplot() +
  geom_sf(data = one_subbasin, fill = NA, color = "black", linewidth = 0.4) +
  geom_sf(
    data = rivers_one,
    aes(color = as.factor(.data$bb_id_prefix)),
    linewidth = 0.5,
    inherit.aes = FALSE
  ) +
  scale_color_discrete(name = "bb_id (first 4 digits)") +
  # Rivers already use color = bb_id; second dam layer needs its own color scale (ggnewscale).
  ggnewscale::new_scale_color() +
  geom_sf(data = dams_current_one, color = "red", size = 2) +
  geom_sf(data = dams_future_other, color = "gray60", size = 2)

# Overlay cascade future dams with distinct colors per dam_id (second scale_color_*).
if (length(cascade_ids) > 0L) {
  p <- p +
    geom_sf(
      data = dams_future_cascade,
      aes(color = as.factor(.data$dam_id)),
      size = 2.5,
      inherit.aes = FALSE
    ) +
    scale_color_manual(
      name = "Future dams (cascade)",
      values = cascade_pal,
      guide = guide_legend(override.aes = list(size = 3))
    )
}

# Minimal map theme: hide axes (geographic context is the data layers).
p +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Sub-basin: trunks by bb_id; future cascade dams by dam_id",
    subtitle = "Red = current dams; gray = future not in cascade"
  )
```
